### 2017-05-08 paper/main.tex
Added reason for objects updating itself

\documentclass{article}
\usepackage[utf8]{inputenc}

% Paper layout
\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.5in}
\oddsidemargin 0.0in 
\evensidemargin .5in
\marginparwidth 0.07 true in
\textheight 7.5 true in
\textwidth 6.5 true in

\PassOptionsToPackage{numbers, compress}{natbib}
\usepackage{url}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{mathtools}
\usepackage{fancyvrb}
\usepackage{array,booktabs,calc,blindtext}
% \usepackage[shortlabels]{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\hmwkTitle}{Extensible Physics Engine}
% \newcommand{\hmwkIssueDate}{22 February 2017}
\newcommand{\hmwkDueDate}{8 May 2017}
\newcommand{\hmwkClass}{6.905/6.945 Final Project Draft}
% \newcommand{\hmwkClassInstructor}{Professor Gerry Sussman}
\newcommand{\hmwkAuthorName}{\textbf{Author1, Author2, Author3}}

\title{
    \textmd{\hmwkClass:\ \hmwkTitle}\\
    \normalsize{\hmwkClassInstructor}\\
    % \small{Issued\ on\ \hmwkIssueDate}\\
    \small{Due\ on\ \hmwkDueDate}
}
\author{\hmwkAuthorName}
\date{}
\begin{document}
\maketitle
\section{Introduction}
motivation: features of physics engine that incorporate ideas from class (e.g. generic operator, hierarchy, propagator, language for language (high--low)), cool visualization of concepts from class

\section{Infrastructure}
generic procedures, ps4 stuff
what we did, how we did it, why we did it

\section{Arithmetic}
We incorporated the generic arithmetic developed in pset 3. We extended the vector-arithmetic to support element-wise operations between scalars and vectors. We did this by coercing scalars to become vectors that contain the same value. For example, \texttt{(+ \#(3 4) 1)} gets transformed into \texttt{(+ \#(3 4) \#(1 1))}.

This extension was made to allow our system to operate on any $n$th dimension vector, and therefore be applicable to any $n$th dimension world representation. The default velocity of an object is set to 0 if it is not specified. Supporting arithmetic operations between a scalar and a vector enables our system to use a single default value 0 that is compatible with any vector instead of limiting our system to the dimension of the default zero vector specified.

The coercion method is 
\begin{verbatim}
(define (coerce a size-reference)
  (if (vector? a)
    a
    (make-vector (vector-length size-reference) a)))
\end{verbatim}
The rest of the infrastructure for generic procedures was adapted from pset 4, but pset 4's generic arithmetic is not compatible with that of pset 3. To fix this issue, we first loaded the relevant pset 3 dependencies, installed a generic arithmetic with vector operations, redefined arithmetic operations to be the ones from that generic arithmetic (such as with \texttt{(define * (access * arith-environment))}), and then loaded the pset 4 generic procedure infrastructure on top.

\section{Physics Engine}
\subsection{Object Types}
There are three types of objects in the system: \texttt{thing}, \texttt{interaction}, and \texttt{world}. \texttt{thing} refers to all physical objects in the system. There is a hierarchy of physical objects that have additional properties. For example, type \texttt{magnet} has an additional property \texttt{charge} and experiences magnetic forces while type \texttt{ball} experiences only gravity. Each \texttt{thing} keeps a list of \texttt{interaction} that is used to update itself. \texttt{interaction} refers to forces exerted on \texttt{thing}s. \texttt{interaction} includes a procedure that calculates the force as well as \texttt{influences}, a list of \texttt{thing}s that contribute to that force. Similar to \texttt{thing}, subtypes of \texttt{interaction}s form a hierarchy. Lastly, \texttt{world} keeps a list of all \texttt{thing}s and an \texttt{update} call to the \texttt{world} calls on each \texttt{thing} to update itself.

(subject to change after restructuring)

\subsection{Updating Physical Objects}
There are mainly two ways positions of physical objects could be updated: 1) The evolver can take in a list of forces and objects and apply forces as appropriate 2) Each object can be registered with a set of forces applicable to it and the evolver can call each object to evolve itself.

Our system adopted the second model where each object updates itself. This model makes it easier to handle different types of forces exerted on specific types of object. Moreover, in the second model, the engine does not need to be notified and updated everytime a new object is added. Instead, update or addition of new \texttt{interaction}s is taken care of at the object level when it is added.

\subsection{Output Display}
Our engine updates each object every timestep and re-renders them on the screen:
\begin{verbatim}
(define (run-engine world steps)
  (reset-graphics)
  (if (> steps 0)
    (begin 
      (for-each 
        (lambda (thing)
            (newline)
            (display (cons (get-name thing) (get-position thing)))  ; for debugging
            (render thing))
        (get-world-all-things world))
      (update-world world)
    (run-engine world (- steps 1)))))
\end{verbatim}

where the \texttt{render} procedure takes in a \texttt{thing}, checks what type of thing it is, and renders the object appropriately.

\section{Demos}
\subsection{Gravity}
% \begin{samepage}
We can easily extend our system with new forces and new objects. We begin by creating an gravity as an \texttt{interaction}:
{\small\begin{verbatim}
(define (make-gravity thing all-things)
  (define (procedure thing influences)
    (sum (map (lambda (influence)
                (let* ((m1 (get-mass thing))
                       (m2 (get-mass influence))
                       (G 6.674e-11)
                       (v (- (get-position influence)  ; vector between influence and thing
                            (get-position thing)))
                       (r (magnitude v))               ; distance between influence and thing
                       (u (/ v r))                     ; unit vector
                       (gmag (* (* G m1 m2)            ; magnitude of gravity
                                (/ 1 (square r)))))
                  (* u gmag)))                        ; gravitational force vector
              influences)))
  (let ((influences (delq thing all-things)))
    (make-interaction gravity? 'gravity procedure influences)))
\end{verbatim}}
% \end{samepage}

\begin{figure}[h!]
  \centering
 \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figs/gravity.png}
  \caption{[\textit{left}] \texttt{earth-moon}; [\textit{center}] \texttt{binary-stars}; [\textit{right}] \texttt{solar-system}}
  \label{figure:gravity}
\end{figure}

We've created several worlds that exhibit gravity (Fig. \ref{figure:gravity}). Once an object is added to the world, any object that has mass will feel gravitational forces with other object that have mass.

% \begin{samepage}
This world contains an earth with a moon orbiting around it.
{\small\begin{verbatim}
(define (earth-moon)
  (define w (make-world "world"))
  (define b1 (make-ball "earth" 30 1e15 #(0 0) #(0 0) "blue"))
  (define b2 (make-ball "moon" 5 1e5 #(100 100) #(-15.361 15.361) "#85929E"))

  (add-mass! b1 w)
  (add-mass! b2 w)
  w)
\end{verbatim}}
% \end{samepage}

% \begin{samepage}
This world contains two equal masses orbiting a common center of mass.
{\small\begin{verbatim}
(define (binary-stars)
  (define w (make-world "world"))
  (define b1 (make-ball "ball1" 5 1e15 #(-100 -100) #(9 -9) "red"))
  (define b2 (make-ball "ball2" 5 1e15 #(100 100) #(-9 9) "green"))

  (add-mass! b1 w)
  (add-mass! b2 w)
  w)
\end{verbatim}}
% \end{samepage}

% \begin{samepage}
This world contains a massive sun and several planets orbitting around the sun.
{\small\begin{verbatim}
(define (solar-system)
  (define w (make-world "world"))
  (define s (make-ball "sun" 30 1e15 #(0 0) #(0 0) "yellow"))
  (define b1 (make-ball "ball1" 5 1e5 #(100 100) #(-15.361 15.361) "blue"))
  (define b2 (make-ball "ball2" 5 1e5 #(110 110) #(-15.361 15.361) "red"))
  (define b3 (make-ball "ball3" 5 1e5 #(120 120) #(-15.361 15.361) "green"))
  (define b4 (make-ball "ball4" 5 1e5 #(130 130) #(-15.361 15.361) "purple"))
  (define b5 (make-ball "ball5" 5 1e5 #(140 140) #(-15.361 15.361) "orange"))
  (define b6 (make-ball "ball6" 5 1e5 #(150 150) #(-15.361 15.361) "gray"))

  (add-mass! s w)
  (add-mass! b1 w)
  (add-mass! b2 w)
  (add-mass! b3 w)
  (add-mass! b4 w)
  (add-mass! b5 w)
  (add-mass! b6 w)
  w)
\end{verbatim}}
% \end{samepage}

\subsection{Magentism}
% \begin{samepage}
Let's add magnetism into the physics engine. 
{\small\begin{verbatim}
(define (make-magnetic-force magnet all-magnets)
  (define (procedure magnet influences)
    (sum (map (lambda (influence)
                (let* ((q1 (get-magnet-charge magnet))
                           (q2 (get-magnet-charge influence))
                           (mu 1.256e-6)  ; permeability of air
                           (v (- (get-position influence)  ; vector between influence and magnet
                             (get-position magnet)))
                           (r (magnitude v))               ; distance between influence and magnet
                           (u (* -1 (/ v r)))              ; unit vector
                           (mmag (* (* mu q1 q2)           ; magnitude of magnetic force
                                (/ 1 (* 4 pi (square r))))))
                  (* u mmag)))                             ; magnetic force vector
              influences)))
  (let ((influences (delq magnet all-magnets)))
    (make-interaction magnetic-force? 'magnetic-force procedure influences)))
\end{verbatim}}
% \end{samepage}

% \begin{samepage}
This is a simple world that two magnets repelling each other.
{\small\begin{verbatim}
(define (magnets-1)
  (define w (make-world "world"))
  (define m1 (make-magnet "magnet1" 1e5 1 #(-30 -30) #(0 0) "red"))
  (define m2 (make-magnet "magnet2" 1e5 1 #(30 30) #(0 0) "red"))

  (add-magnet! m1 w)
  (add-magnet! m2 w)
  w)
\end{verbatim}}
% \end{samepage}

% \begin{samepage}
This world contains two positively charged masses (blue) and two negatively charged masses (red).
{\small\begin{verbatim}
(define (magnets-2)
  (define w (make-world "world"))
  (define m1 (make-magnet "magnet1" 5e5 1 #(-100 -100) #(0 0) "red"))
  (define m2 (make-magnet "magnet2" -5e5 1 #(100 100) #(0 0) "blue"))
  (define m3 (make-magnet "magnet3" 5e5 1 #(-100 100) #(0 0) "red"))
  (define m4 (make-magnet "magnet4" -5e5 1 #(100 -100) #(0 0) "blue"))

  (add-magnet! m1 w)
  (add-magnet! m2 w)
  (add-magnet! m3 w)
  (add-magnet! m4 w)
  w)
\end{verbatim}}
% \end{samepage}

% \begin{samepage}
We can also combine gravity and magnetism together. Here, we created a solar-system-like world where each of the planets repel each other and the sun attracts the planets via gravity and magnetism. As figure \ref{figure:magnetism} (right) shows, this magnetic solar system exhibits wider and diverging orbits compared to figure \ref{figure:gravity} (right), the solar system with only gravity. 
{\small\begin{verbatim}
(define (magnetic-solar-system)
  (define w (make-world "world"))
  (define s (make-magnet "sun" 5e5 1e15 #(0 0) #(0 0) "blue"))
  (define b1 (make-magnet "ball1" -2e7 1e5 #(100 100) #(-15.361 15.361) "red"))
  (define b2 (make-magnet "ball2" -2e7 1e5 #(110 110) #(-15.361 15.361) "red"))
  (define b3 (make-magnet "ball3" -2e7 1e5 #(120 120) #(-15.361 15.361) "red"))
  (define b4 (make-magnet "ball4" -2e7 1e5 #(130 130) #(-15.361 15.361) "red"))
  (define b5 (make-magnet "ball5" -2e7 1e5 #(140 140) #(-15.361 15.361) "red"))
  (define b6 (make-magnet "ball6" -2e7 1e5 #(150 150) #(-15.361 15.361) "red"))

  (add-magnet! s w)
  (add-magnet! b1 w)
  (add-magnet! b2 w)
  (add-magnet! b3 w)
  (add-magnet! b4 w)
  (add-magnet! b5 w)
  (add-magnet! b6 w)
  w)
\end{verbatim}}
% \end{samepage}

\begin{figure}[h!]
  \centering
 \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figs/magnetism.png}
  \caption{[\textit{left}] \texttt{magnets-1}; [\textit{center}] \texttt{magnets-2}; [\textit{right}] \texttt{magnetic-solar-system}}
  \label{figure:magnetism}
\end{figure}

\end{document}